Competing Bandits:\\ The Perils of Exploration under Competition \thanks{This is an extended abstract of \citet{CompetingBandits-merged}, which in turn subsumes conference papers \citet{CompetingBandits-itcs18} and \citet{CompetingBandits-ec19}.}

### 2020-11-25 journal paper/content/sec-intro.tex
Minor edits.

\textbf{(ii)} We dilute the \HardMax agents with a small fraction of ``random agents" who choose a principal uniformly at random. 
\asedit{(One can interpret them as consumers that are not aware of the reputation signals or the principals' strategies.)} We call this model \HardMaxRandom. Then better algorithms help in a big way: a sufficiently better algorithm is guaranteed to win all non-random agents after an initial learning phase. There is a caveat, however: any algorithm can be defeated by interleaving it with the greedy algorithm. This has two undesirable consequences: a better algorithm may sometimes lose in competition, and a Nash equilibrium typically does not exist.

\textbf{(iii)} We further soften the decision rule so that the selection probabilities vary smoothly in terms of principals' Bayesian-expected rewards;
%as a function of the difference between  principals' Bayesian-expected rewards;
We call it \SoftMaxRandom, \asedit{a realistic middle ground between \HardMax agents and random agents.}%
\footnote{\asedit{Alternatively}, one can obtain a \SoftMaxRandom decision rule using a mixture of more ``basic" agent types that follow \HardMax unless the principal's Bayesian-expected rewards are too close to each other.}
\asedit{In the most technical result of the paper, we find that} a better algorithm prevails under much weaker assumptions.
% on what constitutes a better algorithm.
%This is the most technical result of the paper.

>>>>

\textbf{(ii)} We dilute the \HardMax agents with a small fraction of ``random agents" who choose a principal uniformly at random. 
\asedit{(They can be interpreted as consumers that are oblivious to the principals' reputation.)} We call this model \HardMaxRandom. Then better algorithms help in a big way: a sufficiently better algorithm is guaranteed to win all non-random agents after an initial learning phase. There is a caveat, however: any algorithm can be defeated by interleaving it with the greedy algorithm. This has two undesirable consequences: a better algorithm may sometimes lose in competition, and a Nash equilibrium typically does not exist.

\textbf{(iii)} We further soften the decision rule so that the selection probabilities vary smoothly in terms of principals' Bayesian-expected rewards;
%as a function of the difference between  principals' Bayesian-expected rewards;
We call it \SoftMaxRandom, \asedit{a more realistic middle ground between \HardMax agents and random agents.}%
\footnote{\asedit{Alternatively}, one can obtain a \SoftMaxRandom decision rule using a mixture of more ``basic" agent types that follow \HardMax unless the principal's Bayesian-expected rewards are too close to each other.}
\asedit{In the most technical result of the paper, we find that} a better algorithm prevails under much weaker assumptions.
% on what constitutes a better algorithm.
%This is the most technical result of the paper.

### 2020-11-25 journal paper/content/sec-model.tex
Minor edits.

\HardMaxRandom and \SoftMaxRandom can be seen as agent types that, resp., give some chance to the other principal and avoid sharp transitions in their probabilistic response. Alternatively, they can be realized as mixtures over more ``basic" agent types. Indeed, \HardMaxRandom is a mixture of \HardMax and ``random agents" (which choose a principal uniformly at random). One can obtain a \SoftMaxRandom response function using agent types that choose a principal $i$ with a largest reward estimate $\est_i$, unless $|\est_1-\est_2|$ is upper-bounded by some parameter $\theta$, in which case they choose uniformly. Then, we obtain \SoftMaxRandom as a mixture of random agents and these ``$\theta$-\HardMax'' agents, for a suitable distribution over $\theta$. \gaedit{We provide an interpretation of the ``random agents" as reputation-agnostic consumers that are not aware of or do not seek out the reputation signals available to them in the respective choice model variants.}

@@

\xhdr{Competition game.}
Some of our results explicitly study the game between the two principals, termed the \emph{competition game}. We model it as a simultaneous-move game: before the first agent arrives, each principal commits to an MAB algorithm.
Pure strategies in this game correspond to MAB algorithms.%
\footnote{Note that a mixed strategy is also a (randomized) MAB algorithm, \ie also a pure strategy.}
 Principal's utility is defined as the market share, \ie the number of agents that chose this principal. Principals are risk-neutral \gaedit{and aim to maximize their expected utility.}\gadelete{ : they optimize their expected utility.}

>>>>

\HardMaxRandom and \SoftMaxRandom can be seen as agent types that, resp., give some chance to the other principal and avoid sharp transitions in their probabilistic response. Alternatively, they can be realized as mixtures over more ``basic" agent types. Indeed, \HardMaxRandom is a mixture of \HardMax and ``random agents" (which choose a principal uniformly at random). \asedit{The latter can be interpreted as consumers oblivious to principals' reputation, due to the lack of awareness or interest.} One can obtain a \SoftMaxRandom response function using agent types that choose a principal $i$ with a largest reward estimate $\est_i$, unless $|\est_1-\est_2|$ is upper-bounded by some parameter $\theta$, in which case they choose uniformly. Then, we obtain \SoftMaxRandom as a mixture of random agents and these ``$\theta$-\HardMax'' agents, for a suitable distribution over $\theta$.

@@

\xhdr{Competition game.}
Some of our results explicitly study the game between the two principals, termed the \emph{competition game}. We model it as a simultaneous-move game: before the first agent arrives, each principal commits to an MAB algorithm.
Pure strategies in this game correspond to MAB algorithms.%
\footnote{Note that a mixed strategy is also a (randomized) MAB algorithm, \ie also a pure strategy.}
 Principal's utility is defined as the market share, \ie the number of agents that chose this principal. Principals are risk-neutral and aim to maximize their expected utility.
